---
title: "R Notebook"
output: github_document
---
This is a simple R program ot build, train, and test a regression NN on the Boston MASS data set. 

First, load the necessary packages
```{r}
library("neuralnet")
library(MASS)
```

Set the seed for reproducibility. 
```{r}
set.seed(1)
```

Load the data and scale it
```{r}
data = Boston
max_data <- apply(data, 2, max)
min_data <- apply(data, 2, min)
data_scaled <- scale(data,center = min_data, scale = max_data - min_data)
```

split the data using a 70-30 split

```{r}
index = sample(1:nrow(data),round(0.70*nrow(data)))
train_data <- as.data.frame(data_scaled[index,])
test_data <- as.data.frame(data_scaled[-index,])
```

To save time and avoid having to write out all the variables, we get a names vector, and then create a formula. inside the formula so that medv is the response. we then select all the other 'names'/predictors in the names vector and put a '+' so that we can  fully construct the "response ~ var1+var2 +..." format
```{r}
n = names(data)
f = as.formula(paste("medv ~", paste(n[!n %in% "medv"], collapse = " + ")))
```

Create a neural net
```{r}
net_data = neuralnet(f,data=train_data,hidden=10,linear.output=T)
```

plot the neural net
```{r}
plot(net_data)
```

next we are interested to see how well or NN performs. To do this, we first get the predicted medb of testing data using the neural network. 
```{r}
predict_net_test <- compute(net_data,test_data[,1:13])
```


```{r}
predict_net_test_start <- predict_net_test$net.result*(max(data$medv)-min(data$medv))+min(data$medv)
```

```{r}
test_start <- as.data.frame((test_data$medv)*(max(data$medv)-min(data$medv))+min(data$medv))
```

```{r}
MSE.net_data <- sum((test_start - predict_net_test_start)^2)/nrow(test_start)
```



```{r}
Regression_Model <- lm(medv~., data=data)
summary(Regression_Model)
test <- data[-index,]
predict_lm <- predict(Regression_Model,test)
MSE.lm <- sum((predict_lm - test$medv)^2)/nrow(test)
MSE.net_data
MSE.lm

```
